Bifurcation or state tipping: assessing transition type in a model trophic cascade

### 2018-07-13 paper/paper.Rmd
figure

@@

Decreases in adult bass abundance result from mortality ($m_A A$) and harvest ($qA$), and increases result from the maturation of the fraction of $J$ that survive ($s$) mortality (Eq. 1). Planktivorous fish diffuse ($D_F$) between a refuge of size $F_o$ and a foraging arena, and the predation rate of $A$ on $F$ is controlled by $c_{FA}$ (Eq. 2). Growth in $J$ is controlled by the fecundity ($f$) of $A$, but are subject to losses via cannibalization at a rate $c_{JA}$, and interactions with $F$ which are controlled by the rate at which $J$ hide ($h$) in a refuge or become vulnerable ($v$) to predation at a rate $c_{JF}$ as they enter an implicit foraging arena (Eq. 3). Herbivorous zooplankton diffuse between surface waters and a deep water refuge at a rate $D_H$, grow by preying on phytoplankton at rate cHP and converting phytoplankton biomass (phosphorus) by a factor $\alpha$, and decline as they are preyed upon by planktivores at a rate $c_{HF}$ (Eq. 4). Phytoplankton death is due to mortality and sinking at rate m or predation by herbivores at rate cHF, while their growth is controlled by the per capita growth rate ($r_P$), the availability of phosphorus (introduced at rate $L$), and the availability of light at the surface ($I_0$) and the density dependent growth response to phytoplankton in response to light ($\gamma$, Eq. 5).

>>>>

![S-tipping vs B-tipping.  A critical transition in a system with alternative stable states can be driven either by a perturbation to the state, driving the system across the separatrix (top panel), or by a bifurcation, in which a stable point is annihilated (bottom panel)](figures/scheffer-fig.png)

@@

Decreases in adult bass abundance result from mortality ($m_A A$) and harvest ($qA$), and increases result from the maturation of the fraction of $J$ that survive ($s$) mortality (Eq. 1). Planktivorous fish diffuse ($D_F$) between a refuge of size $F_o$ and a foraging arena, and the predation rate of $A$ on $F$ is controlled by $c_{FA}$ (Eq. 2). Growth in $J$ is controlled by the fecundity ($f$) of $A$, but are subject to losses via cannibalization at a rate $c_{JA}$, and interactions with $F$ which are controlled by the rate at which $J$ hide ($h$) in a refuge or become vulnerable ($v$) to predation at a rate $c_{JF}$ as they enter an implicit foraging arena (Eq. 3). Herbivorous zooplankton diffuse between surface waters and a deep water refuge at a rate $D_H$, grow by preying on phytoplankton at rate $c_{HP}$ and converting phytoplankton biomass (phosphorus) by a factor $\alpha$, and decline as they are preyed upon by planktivores at a rate $c_{HF}$ (Eq. 4). Phytoplankton death is due to mortality and sinking at rate m or predation by herbivores at rate cHF, while their growth is controlled by the per capita growth rate ($r_P$), the availability of phosphorus (introduced at rate $L$), and the availability of light at the surface ($I_0$) and the density dependent growth response to phytoplankton in response to light ($\gamma$, Eq. 5).
